Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-28 paper/paper.tex
pass on intro

This paper focuses on rule-based \nids{}, which are very popular in
industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
  intrusion detector.} checks if the network traffic matches a fixed
set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
of Suricata~\cite{suricata}, a popular open-source \nids{}. This
particular rule prescribes a method to capture an Active
Reconnaissance attack for discovering which ports of a publicly
accessible machine are open for connection. Rulesets exist for
Snort~\cite{snort} and Suricata~\cite{suricata}, providing protection
against existing attacks. Unfortunately, attackers actively look for
new mechanisms to exploit systems. For example, Figure~\Fix{plot}
shows the increase in number of rules added to the ``Emerging Threats
Open Ruleset''---a popular public data set of \suri\ rules---in the
past 12 months.

@@

Intuitively, rulesets need to be constantly revised to protect a
system against attackers. Unfortunately, creating these rules is
time-consuming as the language for describing rules is non-trivial
(see Section~\ref{sec:example-suricata-rules}) and it is error-prone
as checking whether the rule captures only malicious traffic is
non-trivial.

@@

This paper proposes \tname{}\footnote{Abbreviation for
  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
technique that uses machine intelligence to automatically synthesize
rules for rule-based \nids. Our goal is to facilitate the creation of
rules and, consequently, harden the protection of \nids\ to network
attacks. \tname{} synthesizes rules from malicious and benign traffic
and from examples of correct rules.

>>>>

This paper focuses on rule-based \nids{}, which are very popular in
industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
  intrusion detector.} checks if the network traffic matches a fixed
set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
of Suricata~\cite{suricata}, a popular open-source \nids{}. This
particular rule prescribes a method to capture an Active
Reconnaissance attack for discovering which ports of a publicly
accessible machine are open for connection (see
Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
and Suricata~\cite{suricata}, providing protection against existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems. For example, Figure~\Fix{plot} shows the increase in
number of rules added to the ``Emerging Threats Open Ruleset''---a
popular public data set of \suri\ rules---in the past 12 months.

@@

Intuitively, rulesets need to be constantly revised to protect a
system against attackers. Unfortunately, creating these rules is
time-consuming and error-prone. The language for describing rules is
non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
correctness of a rule requires data and careful inspection.

@@

This paper proposes \tname{}\footnote{Abbreviation for
  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
technique that uses machine intelligence to automatically synthesize
rules for rule-based \nids. Our goal is to facilitate the creation of
rules and, consequently, help system administrators to harden the
protection of \nids\ to network attacks. \tname{} synthesizes rules
from malicious and benign traffic and from examples of correct rules.
